On-the-fly Algorithms for GKAT Equivalences

### 2025-01-22 main.tex
anonymize authors

\title{On-the-fly Algorithms for GKAT Equivalences
\thanks{Identify applicable funding agency here. If none, delete this.}
}

\author{\IEEEauthorblockN{Cheng Zhang\textsuperscript{\textsection}}
\IEEEauthorblockA{\textit{Department of Computer Science} \\
\textit{University College London}\\
London, United Kingdom \\
0000-0002-8197-6181}
\and
\IEEEauthorblockN{Qiancheng Fu}
\IEEEauthorblockA{\textit{Department of Computer Science} \\
\textit{Boston University}\\
Boston, USA \\
qcfu@bu.edu}
\and
\IEEEauthorblockN{Hang Ji\textsuperscript{\textsection}}
\IEEEauthorblockA{\textit{Department of Computer Science} \\
\textit{Cornell University}\\
Ithaca, USA \\
hj476@cornell.edu}
\and
\IEEEauthorblockN{Ines Santacruz}
\IEEEauthorblockA{\textit{Department of Computer Science} \\
\textit{Boston University}\\
Boston, USA \\
email address or ORCID}
\and
\IEEEauthorblockN{Marco Gaboardi}
\IEEEauthorblockA{\textit{Department of Computer Science} \\
\textit{Boston University}\\
Boston, USA \\
inessdv15@gmail.com}
}

\maketitle
\begingroup\renewcommand\thefootnote{\textsection}
\footnotetext{Work largely performed at Boston University}
\endgroup

@@

\subsection{Evaluation}\label{sec:performance-implementation}
\begin{table}
\footnotesize
\centering
\setlength{\tabcolsep}{3pt}
\begin{tabular}{l r r r r r}
    Benchmark & DV$_{\textsf{BDD}}$ & TC$_{\textsf{BDD}}$ & DV$_{\textsf{SAT}}$ & TC$_{\textsf{SAT}}$ & SK \\
    \hline
    e250b5p10ne    & 0.18   & 0.18   & 0.16 & 0.17 &    5.82 \\
    e250b5p10eq    & 0.19   & 0.18   & 0.15 & 0.18 &    2.83 \\
    e500b5p50ne    & 0.21   & 0.21   & 0.21 & 0.22 &   37.28 \\
    e500b5p50eq    & 0.22   & 0.21   & 0.20 & 0.27 &   14.06 \\
    e1000b10p100ne & 0.26   & 0.28   & 0.34 & 0.38 & timeout \\
    e1000b10p100eq & 0.28   & 0.26   & 0.28 & 0.41 &   77.83 \\
    e2000b20p200ne & 1.32   & 2.60   & 0.68 & 0.79 & timeout \\
    e2000b20p200eq & 1.92   & 2.95   & 0.41 & 1.00 & timeout \\
    e3000b30p200ne & 1.57   & 24.86  & 1.30 & 1.60 & timeout \\
    e3000b30p200eq & 10.82  & 22.83  & 0.66 & 1.54 & timeout \\
    degenerate     & 99.48  & 228.16 & 0.24 & 0.30 & timeout
\end{tabular}
\caption{Total Time Usage (seconds)}\label{tab:benchmark-time}
\end{table}
\begin{table}
\footnotesize
\centering
\setlength{\tabcolsep}{3pt}
\begin{tabular}{l r r r r r}
    Benchmark & DV$_{\textsf{BDD}}$ & TC$_{\textsf{BDD}}$ & DV$_{\textsf{SAT}}$ & TC$_{\textsf{SAT}}$ & SK \\
    \hline
    e250b5p10ne    & 15.17  & 14.54   & 7.07  & 7.01  &  114.06 \\
    e250b5p10eq    & 15.69  & 14.69   & 7.06  & 7.29  &  100.48 \\
    e500b5p50ne    & 15.99  & 15.33   & 7.02  & 7.29  &  524.89 \\
    e500b5p50eq    & 16.85  & 15.26   & 6.99  & 7.01  &  546.91 \\
    e1000b10p100ne & 17.51  & 21.31   & 7.78  & 7.50  & timeout \\
    e1000b10p100eq & 19.52  & 17.48   & 8.17  & 7.04  & 5822.66 \\
    e2000b20p200ne & 237.17 & 280.70  & 12.95 & 11.43 & timeout \\
    e2000b20p200eq & 107.61 & 102.18  & 12.76 & 13.25 & timeout \\
    e3000b30p200ne & 112.06 & 1232.24 & 21.00 & 19.30 & timeout \\
    e3000b30p200eq & 245.23 & 228.86  & 21.18 & 17.64 & timeout \\
    degenerate     & 632.70 & 1225.24 & 19.13 & 17.94 & timeout
\end{tabular}
\caption{Peak Memory Usage (megabytes)}\label{tab:benchmark-memory}
\vspace{-1em}
\end{table}

Due to the fact that our symbolic algorithms are generic regarding the choice of solvers for boolean satisfiability, we compare their performance using BDD to using a SAT solver (MiniSat~\cite{een_MINISAT_2004}). Our experiments are performed on a laptop with an Apple M4 Pro CPU and 24 GB RAM. \Cref{tab:benchmark-time,tab:benchmark-memory} present the experimental results of equivalence checking based on derivatives (DV), Thompson's construction (TC) and SymKAT~\cite{pous_SymbolicAlgorithmsLanguage_2015} (SK). Note that SymKAT uses BDD internally for checking boolean satisfiability. Each benchmark consists of 50 expression pairs with properties described by the benchmark's name. For example, expressions in `e250b5p10eq' have approximately 250 primitive actions in total (e250), a maximum boolean expression size of 5 (b5), 10 possible unique primitive tests (p10) and are known to be equivalent (eq). Benchmarks with the suffix `ne' have expression pairs that are known to be non-equivalent.

>>>>

\title{On-the-fly Algorithms for GKAT Equivalences
% \thanks{Identify applicable funding agency here. If none, delete this.}
}

% anonymize authors
%
% \author{\IEEEauthorblockN{Cheng Zhang\textsuperscript{\textsection}}
% \IEEEauthorblockA{\textit{Department of Computer Science} \\
% \textit{University College London}\\
% London, United Kingdom \\
% 0000-0002-8197-6181}
% \and
% \IEEEauthorblockN{Qiancheng Fu}
% \IEEEauthorblockA{\textit{Department of Computer Science} \\
% \textit{Boston University}\\
% Boston, USA \\
% qcfu@bu.edu}
% \and
% \IEEEauthorblockN{Hang Ji\textsuperscript{\textsection}}
% \IEEEauthorblockA{\textit{Department of Computer Science} \\
% \textit{Cornell University}\\
% Ithaca, USA \\
% hj476@cornell.edu}
% \and
% \IEEEauthorblockN{Ines Santacruz}
% \IEEEauthorblockA{\textit{Department of Computer Science} \\
% \textit{Boston University}\\
% Boston, USA \\
% inessdv15@gmail.com}
% \and
% \IEEEauthorblockN{Marco Gaboardi}
% \IEEEauthorblockA{\textit{Department of Computer Science} \\
% \textit{Boston University}\\
% Boston, USA \\
% gaboardi@bu.edu}
% }

\maketitle
% \begingroup\renewcommand\thefootnote{\textsection}
% \footnotetext{Work largely performed at Boston University}
% \endgroup

@@

\subsection{Evaluation}\label{sec:performance-implementation}
Due to the fact that our symbolic algorithms are generic regarding the choice of solvers for boolean satisfiability, we compare their performance using BDD to using a SAT solver (MiniSat~\cite{een_MINISAT_2004}). Our experiments are performed on a laptop with an Apple M4 Pro CPU and 24 GB RAM. \Cref{tab:benchmark-time,tab:benchmark-memory} present the experimental results of equivalence checking based on derivatives (DV), Thompson's construction (TC) and SymKAT~\cite{pous_SymbolicAlgorithmsLanguage_2015} (SK). Note that SymKAT uses BDD internally for checking boolean satisfiability. Each benchmark consists of 50 expression pairs with properties described by the benchmark's name. For example, expressions in `e250b5p10eq' have approximately 250 primitive actions in total (e250), a maximum boolean expression size of 5 (b5), 10 possible unique primitive tests (p10) and are known to be equivalent (eq). Benchmarks with the suffix `ne' have expression pairs that are known to be non-equivalent.

\begin{table}
\footnotesize
\centering
\setlength{\tabcolsep}{3pt}
\begin{tabular}{l r r r r r}
    Benchmark & DV$_{\textsf{BDD}}$ & TC$_{\textsf{BDD}}$ & DV$_{\textsf{SAT}}$ & TC$_{\textsf{SAT}}$ & SK \\
    \hline
    e250b5p10ne    & 0.18   & 0.18   & 0.16 & 0.17 &    5.82 \\
    e250b5p10eq    & 0.19   & 0.18   & 0.15 & 0.18 &    2.83 \\
    e500b5p50ne    & 0.21   & 0.21   & 0.21 & 0.22 &   37.28 \\
    e500b5p50eq    & 0.22   & 0.21   & 0.20 & 0.27 &   14.06 \\
    e1000b10p100ne & 0.26   & 0.28   & 0.34 & 0.38 & timeout \\
    e1000b10p100eq & 0.28   & 0.26   & 0.28 & 0.41 &   77.83 \\
    e2000b20p200ne & 1.32   & 2.60   & 0.68 & 0.79 & timeout \\
    e2000b20p200eq & 1.92   & 2.95   & 0.41 & 1.00 & timeout \\
    e3000b30p200ne & 1.57   & 24.86  & 1.30 & 1.60 & timeout \\
    e3000b30p200eq & 10.82  & 22.83  & 0.66 & 1.54 & timeout \\
    degenerate     & 99.48  & 228.16 & 0.24 & 0.30 & timeout
\end{tabular}
\caption{Total Time Usage (seconds)}\label{tab:benchmark-time}
\end{table}
\begin{table}
\footnotesize
\centering
\setlength{\tabcolsep}{3pt}
\begin{tabular}{l r r r r r}
    Benchmark & DV$_{\textsf{BDD}}$ & TC$_{\textsf{BDD}}$ & DV$_{\textsf{SAT}}$ & TC$_{\textsf{SAT}}$ & SK \\
    \hline
    e250b5p10ne    & 15.17  & 14.54   & 7.07  & 7.01  &  114.06 \\
    e250b5p10eq    & 15.69  & 14.69   & 7.06  & 7.29  &  100.48 \\
    e500b5p50ne    & 15.99  & 15.33   & 7.02  & 7.29  &  524.89 \\
    e500b5p50eq    & 16.85  & 15.26   & 6.99  & 7.01  &  546.91 \\
    e1000b10p100ne & 17.51  & 21.31   & 7.78  & 7.50  & timeout \\
    e1000b10p100eq & 19.52  & 17.48   & 8.17  & 7.04  & 5822.66 \\
    e2000b20p200ne & 237.17 & 280.70  & 12.95 & 11.43 & timeout \\
    e2000b20p200eq & 107.61 & 102.18  & 12.76 & 13.25 & timeout \\
    e3000b30p200ne & 112.06 & 1232.24 & 21.00 & 19.30 & timeout \\
    e3000b30p200eq & 245.23 & 228.86  & 21.18 & 17.64 & timeout \\
    degenerate     & 632.70 & 1225.24 & 19.13 & 17.94 & timeout
\end{tabular}
\caption{Peak Memory Usage (megabytes)}\label{tab:benchmark-memory}
\vspace{-1em}
\end{table}
